Negative Perceptions of Outsourcing to Artificial Intelligence

### 2025-04-28 appendix.qmd
First draft of manuscript

# Supplementary Materials {.unnumbered .unlisted}

\setcounter{page}{1}
\begin{center}
Negative Perceptions of Outsourcing to Artificial Intelligence \\
Scott Claessens, Pierce Veitch, and Jim Everett \\
School of Psychology, University of Kent
\end{center}
\newpage
\tableofcontents
\newpage

@@

### Statistical analysis

@@

We presented participants with six scenarios (see @suppfig-scenarios-pilotstudy2 
for examples). Each scenario described a person completing a task, such as 
writing computer code or writing a love letter. The six tasks were randomly 
drawn from a larger set of 20 tasks (see @supptbl-tasks for the full list of 
tasks). For each scenario, we told participants:

@@

@@

::: {#suppfig-tasks-correlations}

```{r tasks-correlations, fig.height=4, fig.width=5}
tar_read(plot_task_cors)
```

Model-estimated task-specific correlations between all six questions. Values are 
posterior median correlations. A positive correlation indicates that tasks that 
are rated highly on one question tend to be rated highly on another question.

@@

::: {#suppfig-tasks-social}

```{r tasks-social, fig.height=4, fig.width=6}
tar_read(plot_task_rank_social)
```

Model-estimated means for the question "Is this a social task?" across all 20 
tasks. Grey points represent participant responses to the question, jittered for 
easier viewing. Black points are estimated means from the fitted model, pooling 
over participants. Black points and line ranges represent posterior medians and 
95% credible intervals, respectively.

@@

::: {#suppfig-tasks-socialskills}

```{r tasks-socialskills, fig.height=4, fig.width=6}
tar_read(plot_task_rank_socialskills)
```

Model-estimated means for the question "Does this task require social skills?" 
across all 20 tasks. Grey points represent participant responses to the 
question, jittered for easier viewing. Black points are estimated means from the 
fitted model, pooling over participants. Black points and line ranges represent 
posterior medians and 95% credible intervals, respectively.

@@

::: {#suppfig-tasks-impactothers}

```{r tasks-impactothers, fig.height=4, fig.width=6}
tar_read(plot_task_rank_impactothers)
```

Model-estimated means for the question "Does this task impact other people?" 
across all 20 tasks. Grey points represent participant responses to the 
question, jittered for easier viewing. Black points are estimated means from the 
fitted model, pooling over participants. Black points and line ranges represent 
posterior medians and 95% credible intervals, respectively.

@@

::: {#suppfig-tasks-consequences}

```{r tasks-consequences, fig.height=4, fig.width=6}
tar_read(plot_task_rank_consequences)
```

Model-estimated means for the question "How important are the consequences of 
this task?" across all 20 tasks. Grey points represent participant responses to 
the question, jittered for easier viewing. Black points are estimated means from 
the fitted model, pooling over participants. Black points and line ranges 
represent posterior medians and 95% credible intervals, respectively.

@@

::: {#suppfig-tasks-intrinsiceffort}

```{r tasks-intrinsiceffort, fig.height=4, fig.width=6}
tar_read(plot_task_rank_intrinsiceffort)
```

Model-estimated means for the question "How important is it that effort goes 
into this task?" across all 20 tasks. Grey points represent participant 
responses to the question, jittered for easier viewing. Black points are 
estimated means from the fitted model, pooling over participants. Black points 
and line ranges represent posterior medians and 95% credible intervals, 
respectively.

@@

::: {#suppfig-tasks-extrinsiceffort}

```{r tasks-extrinsiceffort, fig.height=4, fig.width=6}
tar_read(plot_task_rank_extrinsiceffort)
```

Model-estimated means for the question "How important is it that others see the 
effort that goes into this task?" across all 20 tasks. Grey points represent 
participant responses to the question, jittered for easier viewing. Black points 
are estimated means from the fitted model, pooling over participants. Black 
points and line ranges represent posterior medians and 95% credible intervals, 
respectively.

@@

::: {#suppfig-chatgpt-pilotstudy2}

```{r chatgpt-pilotstudy2, fig.height=5, fig.width=6}
tar_read(plot_chatgpt_study1)
```

@@

::: {#suppfig-scenarios-pilotstudy2}

![](images/scenarios_pilotstudy2.png)

Examples of the scenarios presented to participants in the second pilot study.

::: {#suppfig-treatments-pilotstudy2}

```{r treatments-pilotstudy2, fig.height=4, fig.width=6}
tar_read(plot_treatments_study1)
```

@@

::: {#suppfig-treatments-by-task-pilotstudy2}

```{r treatments-by-task-pilotstudy2, fig.height=5, fig.width=8.5}
tar_read(plot_treatments_tasks_study1)
```

@@

::: {#suppfig-interactions-pilotstudy2}

```{r interactions-pilotstudy2, fig.height=5, fig.width=6.6}
tar_read(plot_interactions_study1)
```

The impact of task-specific features (e.g., being a social task) on the causal
effects of outsourcing to AI (red) and humans (blue) compared to the control
condition. The y-axis reflects the estimated differences between the
experimental conditions and the control condition (dashed line) on a 7-point 
Likert scale. Lines and shaded areas represent posterior medians and 95%
credible intervals, respectively. The patterns indicate, for example, more 
negative effects of outsourcing on character evaluations for tasks that are 
rated as more social.

@@

::: {#suppfig-interaction-pars-pilotstudy2}

```{r interaction-pars-pilotstudy2, fig.height=5.5, fig.width=5.5}
tar_read(plot_interaction_parameters_study1)
```

Interaction parameters from models including task-specific features as
moderators of the causal effects of AI outsourcing (red) and human outsourcing 
(blue) compared to the control condition. Points and line ranges represent 
posterior medians and 95% credible intervals, respectively.

@@

::: {#supptbl-tasks tbl-pos="H" layout-ncol="1"}

```{r tasks}
tar_read(table_tasks) %>%
  kable(
    format = "latex",
    escape = FALSE,
    booktabs = TRUE,
    linesep = "", # to prevent space every five rows
    align = "lccccc"
  ) %>%
  kable_styling(font_size = 10)
```

@@

::: {#supptbl-treatment-diffs-pilotstudy2 tbl-pos="H" layout-ncol="1"}

```{r treatment-diffs-pilotstudy2}
tar_read(table_treatment_diffs_study1) %>%
  kable(
    format = "latex",
    booktabs = TRUE,
    linesep = ""
  ) %>%
  kable_styling(font_size = 9) %>%
  add_header_above(header = c(" " = 1, "Response" = 5))
```

>>>>

# Supplementary Materials

\setcounter{page}{1}
\begin{center}
Negative Perceptions of Outsourcing to Artificial Intelligence \\
Scott Claessens, Pierce Veitch, and Jim Everett \\
School of Psychology, University of Kent
\end{center}
\newpage

@@

### Statistical Analysis

@@

We presented participants with six scenarios. Each scenario described a person 
completing a task, such as writing computer code or writing a love letter. The 
six tasks were randomly drawn from a larger set of 20 tasks (see @supptbl-tasks
for the full list of tasks). For each scenario, we told participants:

@@

# Vignette Wording in Study 5

:::{.NoIndent}
We presented participants in Study 5 with the following vignette text:
:::

> Adam has been dating his partner for almost a year, and Valentine's Day is 
coming up. He knows that many people exchange a card on Valentine's Day 
containing a love letter to their partner, and he decides to send a love letter
to his partner too.
>
> This year, Adam has been closely following developments in technology and has
read of people using AI tools like ChatGPT for things like this, either using it
to help with writing or getting AI to do the task completely.

:::{.NoIndent}
This was followed by the manipulation text (see main text). Participants were 
then presented with the love letter that Adam ostensibly wrote, which was held 
constant across conditions:
:::

> Happy Valentine's Day, my love.
>
> I don't think I tell you enough just how much you mean to me. Being with you 
feels like breathing a little easier, like the world is a bit softer just 
because you're in it. You make the everyday feel special, and somehow you always
know how to calm my nerves or make me laugh at just the right moment. I feel
like myself with you -- maybe even a better version of myself -- and that's such
a rare and beautiful thing.
>
> I'm so grateful for you -- for the way you listen, the way you love, the way
you show up, even in the small ways. I hope you know that no matter what, I'm
always in your corner. I can't wait to keep making memories together, whether
we're off on some adventure or just curled up on the couch. I love you more than
I can really put into words, but I promise I'll spend every day trying.
>
> Yours,
>
> Adam

@@

::: {#suppfig-chatgpt-study1 suppfig-pos="H"}

```{r, fig.height=5, fig.width=6}
tar_read(plot_chatgpt_study2)
```

Responses to the questions about ChatGPT in Study 1.

@@

::: {#suppfig-interactions-study1 suppfig-pos="H"}

```{r, fig.height=6, fig.width=7}
tar_read(plot_interactions_study2)
```

The impact of task-specific features (e.g., being a social task) on the causal
effects of outsourcing to AI compared to the control condition in Study 1. The 
y-axis reflects the estimated differences between the experimental conditions 
and the control condition (dashed line) on a 7-point Likert scale. Lines and 
shaded areas represent posterior medians and 95% credible intervals, 
respectively. The patterns indicate, for example, more negative effects of 
outsourcing on character evaluations for tasks that are rated as more social.

@@

::: {#suppfig-interaction-pars-study1 suppfig-pos="H"}

```{r, fig.height=6, fig.width=6}
tar_read(plot_interaction_parameters_study2)
```

Interaction parameters from models including task-specific features as
moderators of the causal effects of AI outsourcing compared to the control 
condition in Study 1. Points and line ranges represent posterior medians and 95% 
credible intervals, respectively.

@@

::: {#suppfig-chatgpt-study2 suppfig-pos="H"}

```{r, fig.height=5, fig.width=6}
tar_read(plot_chatgpt_study3)
```

Responses to the questions about ChatGPT in Study 2.

@@

::: {#suppfig-treatments-person-study2 suppfig-pos="H"}

```{r, fig.height=4, fig.width=6}
tar_read(plot_treatments_person_study3)
```

Character evaluations in Study 2. Participants in the control condition, the 
tool outsourcing condition, and the full outsourcing condition evaluated the 
"other participant" on (a) competence, (b) warmth, (c) morality, (d) laziness, 
and (e) trustworthiness. Jittered points represent participant responses to the
questions, split by whether the writing task was a non-social task (red) or a 
social task (blue). Point ranges are estimated marginal means from the fitted 
model, pooling over essay answers. Points and line ranges represent posterior 
medians and 95% credible intervals, respectively.

@@

::: {#suppfig-chatgpt-study3 suppfig-pos="H"}

```{r, fig.height=5, fig.width=6}
tar_read(plot_chatgpt_study4)
```

Responses to the questions about ChatGPT in Study 3.

@@

::: {#suppfig-treatments-tasks-study3 suppfig-pos="H"}

```{r, fig.height=5, fig.width=8.5}
tar_read(plot_treatments_tasks_study4)
```

Variation in the effects of outsourcing across tasks in Study 3. Tasks are 
ordered from most social (top) to least social (bottom) according to ratings 
from a pilot study. Point ranges are differences in marginal means on a 7-point
Likert scale for the low effort conditions (red) and high effort conditions 
(blue) compared to the control condition. Upper panels refer to the standard LLM
conditions, and lower panels refer to the personalised LLM conditions. Points 
and ranges represent posterior medians and 95% credible intervals, respectively.

@@

::: {#suppfig-interactions-study3 suppfig-pos="H"}

```{r, fig.height=6, fig.width=7}
tar_read(plot_interactions_study4)
```

The impact of task-specific features (e.g., being a social task) on the causal
effects of outsourcing to AI compared to the control condition in Study 3. The 
y-axis reflects the estimated differences between the experimental conditions 
and the control condition (dashed line) on a 7-point Likert scale. Lines and 
shaded areas represent posterior medians and 95% credible intervals, 
respectively. The patterns indicate, for example, more negative effects of 
outsourcing on character evaluations for tasks that are rated as more social.

::: {#suppfig-interaction-pars-study3 suppfig-pos="H"}

```{r, fig.height=6, fig.width=6}
tar_read(plot_interaction_parameters_study4)
```

Interaction parameters from models including task-specific features as
moderators of the causal effects of AI outsourcing compared to the control 
condition in Study 3. Points and line ranges represent posterior medians and 95% 
credible intervals, respectively.

::: {#suppfig-chatgpt-study4 suppfig-pos="H"}

```{r, fig.height=5, fig.width=6}
tar_read(plot_chatgpt_study5)
```

Responses to the questions about ChatGPT in Study 4.

::: {#suppfig-treatments-tasks-study4 suppfig-pos="H"}

```{r, fig.height=4, fig.width=8.5}
tar_read(plot_treatments_tasks_study5)
```

Variation in the effects of outsourcing across tasks in Study 4. Tasks are 
ordered from most social (top) to least social (bottom) according to ratings 
from a pilot study. Point ranges are differences in marginal means on a 7-point
Likert scale for the bad reason conditions (red) and good reason conditions 
(blue) compared to the control condition. Upper panels refer to the tool 
outsourcing conditions, and lower panels refer to the full outsourcing 
conditions. Points and ranges represent posterior medians and 95% credible 
intervals, respectively.

::: {#suppfig-interactions-study4 suppfig-pos="H"}

```{r, fig.height=6, fig.width=7}
tar_read(plot_interactions_study5)
```

The impact of task-specific features (e.g., being a social task) on the causal
effects of outsourcing to AI compared to the control condition in Study 4. The 
y-axis reflects the estimated differences between the experimental conditions 
and the control condition (dashed line) on a 7-point Likert scale. Lines and 
shaded areas represent posterior medians and 95% credible intervals, 
respectively. The patterns indicate, for example, more negative effects of 
outsourcing on character evaluations for tasks that are rated as more social.

::: {#suppfig-interaction-pars-study4 suppfig-pos="H"}

```{r, fig.height=6, fig.width=6}
tar_read(plot_interaction_parameters_study5)
```

Interaction parameters from models including task-specific features as
moderators of the causal effects of AI outsourcing compared to the control 
condition in Study 4. Points and line ranges represent posterior medians and 95% 
credible intervals, respectively.

::: {#suppfig-chatgpt-study5 suppfig-pos="H"}

```{r, fig.height=5, fig.width=6}
tar_read(plot_chatgpt_study6)
```

Responses to the questions about ChatGPT in Study 5.

::: {#suppfig-tasks-correlations suppfig-pos="H"}

```{r, fig.height=4, fig.width=5}
tar_read(plot_task_cors)
```

Model-estimated task-specific correlations between all six questions in the 
first pilot study. Values are posterior median correlations. A positive 
correlation indicates that tasks that are rated highly on one question tend to 
be rated highly on another question.

::: {#suppfig-tasks-social suppfig-pos="H"}

```{r, fig.height=4, fig.width=6}
tar_read(plot_task_rank_social)
```

Model-estimated means for the question "Is this a social task?" across all 20 
tasks in the first pilot study. Grey points represent participant responses to 
the question, jittered for easier viewing. Black points are estimated means from
the fitted model, pooling over participants. Black points and line ranges 
represent posterior medians and 95% credible intervals, respectively.

::: {#suppfig-tasks-socialskills suppfig-pos="H"}

```{r, fig.height=4, fig.width=6}
tar_read(plot_task_rank_socialskills)
```

Model-estimated means for the question "Does this task require social skills?" 
across all 20 tasks in the first pilot study. Grey points represent participant 
responses to the question, jittered for easier viewing. Black points are 
estimated means from the fitted model, pooling over participants. Black points 
and line ranges represent posterior medians and 95% credible intervals, 
respectively.

::: {#suppfig-tasks-impactothers suppfig-pos="H"}

```{r, fig.height=4, fig.width=6}
tar_read(plot_task_rank_impactothers)
```

Model-estimated means for the question "Does this task impact other people?" 
across all 20 tasks in the first pilot study. Grey points represent participant 
responses to the question, jittered for easier viewing. Black points are 
estimated means from the fitted model, pooling over participants. Black points 
and line ranges represent posterior medians and 95% credible intervals,
respectively.

::: {#suppfig-tasks-consequences suppfig-pos="H"}

```{r, fig.height=4, fig.width=6}
tar_read(plot_task_rank_consequences)
```

Model-estimated means for the question "How important are the consequences of 
this task?" across all 20 tasks in the first pilot study. Grey points represent 
participant responses to the question, jittered for easier viewing. Black points
are estimated means from the fitted model, pooling over participants. Black
points and line ranges represent posterior medians and 95% credible intervals, 
respectively.

::: {#suppfig-tasks-intrinsiceffort suppfig-pos="H"}

```{r, fig.height=4, fig.width=6}
tar_read(plot_task_rank_intrinsiceffort)
```

Model-estimated means for the question "How important is it that effort goes 
into this task?" across all 20 tasks in the first pilot study. Grey points 
represent participant responses to the question, jittered for easier viewing.
Black points are estimated means from the fitted model, pooling over
participants. Black points and line ranges represent posterior medians and 95% 
credible intervals, respectively.

::: {#suppfig-tasks-extrinsiceffort suppfig-pos="H"}

```{r, fig.height=4, fig.width=6}
tar_read(plot_task_rank_extrinsiceffort)
```

Model-estimated means for the question "How important is it that others see the 
effort that goes into this task?" across all 20 tasks in the first pilot study.
Grey points represent participant responses to the question, jittered for easier 
viewing. Black points are estimated means from the fitted model, pooling over 
participants. Black points and line ranges represent posterior medians and 95% 
credible intervals, respectively.

::: {#suppfig-chatgpt-pilotstudy2 suppfig-pos="H"}

```{r, fig.height=5, fig.width=6}
tar_read(plot_chatgpt_study1)
```

@@

::: {#suppfig-treatments-pilotstudy2 suppfig-pos="H"}

```{r, fig.height=4, fig.width=6}
tar_read(plot_treatments_study1)
```

@@

::: {#suppfig-treatments-by-task-pilotstudy2 suppfig-pos="H"}

```{r, fig.height=5, fig.width=8.5}
tar_read(plot_treatments_tasks_study1)
```

@@

::: {#suppfig-interactions-pilotstudy2 suppfig-pos="H"}

```{r, fig.height=5, fig.width=6.6}
tar_read(plot_interactions_study1)
```

The impact of task-specific features (e.g., being a social task) on the causal
effects of outsourcing to AI (red) and humans (blue) compared to the control
condition in the second pilot study. The y-axis reflects the estimated 
differences between the experimental conditions and the control condition 
(dashed line) on a 7-point Likert scale. Lines and shaded areas represent 
posterior medians and 95% credible intervals, respectively. The patterns 
indicate, for example, more negative effects of outsourcing on character 
evaluations for tasks that are rated as more social.

@@

::: {#suppfig-interaction-pars-pilotstudy2 suppfig-pos="H"}

```{r, fig.height=5.5, fig.width=5.5}
tar_read(plot_interaction_parameters_study1)
```

Interaction parameters from models including task-specific features as
moderators of the causal effects of AI outsourcing (red) and human outsourcing 
(blue) compared to the control condition in the second pilot study. Points and 
line ranges represent posterior medians and 95% credible intervals,
respectively.

@@

::: {#supptbl-tasks supptbl-pos="H" layout-ncol="1"}

```{r}
tar_read(table_tasks) %>%
  kable(
    format = "latex",
    escape = FALSE,
    booktabs = TRUE,
    linesep = "",
    align = "lccccc"
  ) %>%
  kable_styling(font_size = 10)
```

@@

::: {#supptbl-essay-answers-social-study2 supptbl-pos="H" layout-ncol="1"}

```{r}
table <- 
  tar_read(table_essay_answers_study3) %>%
  slice(1:3) %>%
  kable(
    format = "latex",
    booktabs = TRUE,
    linesep = ""
  ) %>%
  kable_styling(font_size = 10) %>%
  column_spec(1, width = "2cm") %>%
  column_spec(2, width = "14cm")

# table single spaced
paste(
  "\\begingroup",
  "\\linespread{1}\\selectfont",
  table,
  "\\endgroup",
  sep = "\n"
  ) %>%
  knitr::asis_output()
```

Pre-generated essay answers to the social prompt in Study 2.

::: {#supptbl-essay-answers-nonsocial-study2 supptbl-pos="H" layout-ncol="1"}

```{r}
table <- 
  tar_read(table_essay_answers_study3) %>%
  slice(4:6) %>%
  kable(
    format = "latex",
    booktabs = TRUE,
    linesep = ""
  ) %>%
  kable_styling(font_size = 10) %>%
  column_spec(1, width = "2cm") %>%
  column_spec(2, width = "14cm")

# table single spaced
paste(
  "\\begingroup",
  "\\linespread{1}\\selectfont",
  table,
  "\\endgroup",
  sep = "\n"
  ) %>%
  knitr::asis_output()
```

Pre-generated essay answers to the non-social prompt in Study 2.

::: {#supptbl-essay-comprehension-study2 supptbl-pos="H" layout-ncol="1"}

```{r}
tar_read(table_essay_comprehension_study3) %>%
  kable(
    format = "latex",
    booktabs = TRUE,
    linesep = ""
  ) %>%
  kable_styling(font_size = 9)
```

Reading times and comprehension rates for the essay answers in Study 2. Expected
reading times were calculated based on an estimated reading speed of 275 words
per minute. Comprehension rates are the percentage of participants who answered
the comprehension question correctly.

::: {#supptbl-manipulation-check-study2 supptbl-pos="H" layout-ncol="1"}

```{r}
tar_read(table_manipulation_check_study3) %>%
  kable(
    format = "latex",
    booktabs = TRUE,
    linesep = ""
  ) %>%
  kable_styling(font_size = 11)
```

Percentage of participants in Study 2 who passed the manipulation check and 
reported that they believed the manipulation, split by condition.

::: {#supptbl-treatment-diffs-person-study2 supptbl-pos="H" layout-ncol="1"}

```{r}
tar_read(table_treatment_diffs_effects_study3) %>%
  dplyr::select(c(1, 7:11)) %>%
  kable(
    format = "latex",
    booktabs = TRUE,
    linesep = ""
  ) %>%
  kable_styling(font_size = 7.5) %>%
  add_header_above(header = c(" " = 1, "Response" = 5)) %>%
  pack_rows(
    index = c(
      "Effect of outsourcing type" = 6,
      "Effect of task type" = 3,
      "Interaction effect" = 3
      )
    ) %>%
  pack_rows("\u00A0\u00A0Task type = Social", 1, 3, indent = FALSE) %>%
  pack_rows("\u00A0\u00A0Task type = Non-social", 4, 6, indent = FALSE)
```

Pairwise contrasts for character evaluations in Study 2. Numbers reflect 
differences in marginal means on a 7-point Likert scale. Estimates are pooled 
over essay answers. The bottom rows represent the interactions between 
outsourcing type and task type. Main numbers are posterior medians, numbers in 
the square brackets are 95% credible intervals.

::: {.landscape}

::: {#supptbl-text-analysis-study5 supptbl-pos="H" layout-ncol="1"}

```{r}
tar_read(table_text_analysis_study6) %>%
  kable(
    format = "latex",
    booktabs = TRUE,
    linesep = ""
  ) %>%
  kable_styling(font_size = 8)
```

Pairwise comparisons of word frequencies between conditions. LL = log 
likelihood; %DIFF =

::: {#supptbl-treatment-diffs-pilotstudy2 supptbl-pos="H" layout-ncol="1"}

```{r}
tar_read(table_treatment_diffs_study1) %>%
  kable(
    format = "latex",
    booktabs = TRUE,
    linesep = ""
  ) %>%
  kable_styling(font_size = 9) %>%
  add_header_above(header = c(" " = 1, "Response" = 5))
```
